---
title: "Test a Perceptual Phenomenon"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
knitr::opts_chunk$set(echo = TRUE)
```

#### By Hardeep Johar

1.)   
Independent variable: congruent vs. incongruent words.   
Dependent variable: Time it takes to name the ink colors in equally-sized lists.

2a.)   
Null hypothesis (H0): There is no difference in population means of response   
time for incongruent and congruent scenarios, which can be mathematically  
represented as:   
H0: uC = uI where uC is the population mean of response time for incongruent   
words and uI is the population mean of reponse time for congruent words.

Alternative hypothesis (HA): Population mean of response time for congruent   
scenarios is less than population mean of response time for incongruent   
scenarios, which can be mathematically represented as:   
HA: uC < uI where uC is the population mean of response time for incongruent   
words and uI is the population mean of reponse time for congruent words.

2b.)   
I expect to perform a one tailed paired t-test.  One tailed because we are   
concerned with how much more time and not how much less time - so the test    
would only need to be one directional.  Also because we do not know the   
population parameters a t-test would be the most apprpriate.

A Z-test would require we have greater than 30 samples and we know the    
population mean (sigma).  Neither is true in this case so we cannot use a   
Z-test.

An F-test would require 3 or more population means which we do not have so   
we cannot use an F-test.

We are trying to determine the following: based on these samples,   
can we conclude that there is no difference in population means of response   
time for incongruent and congruent scenarios at a given confidence level.


```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
persep <- read.csv('stroopdata.csv')
persep2 <- gather(persep, key=Group, value=Time)
```

3.)  
Congruent dataset standard deviation is 3.559.  
Below shows the measure of center and variability.

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(persep$Congruent)
```

Incongruent dataset standard deviation is 4.797.   
Below shows the measure of center and variability.

```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(persep$Incongruent)
```

4.)   

```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(persep2, aes(y = Time, x = Group, col=Group)) + 
              geom_jitter(position = position_jitter(height = 0, width=0.1))+
              geom_boxplot(alpha = 0, color = "black")
```

This chart is a visualization of the measures of variability and center shown   
above.  It shows that in most cases, congruent words take less time to read    
through than Incongruent words do.

5.)   
At a confidence level of 95% (alpha = 0.05) on a 1 tailed T test with 23 dF:   
T-Critical = 1.714   
T-Statistic = -8.021   
We would reject the null hypothesis   
The results did match up with my expectations that incongruent word lists will   
take more time to read through than iongruent word lists.    

6.)   
What is responsible for the effects is the brain getting tricked into reading   
the words instead of distinguishing the color.

An example of this could be asking a group of people to choose a shoe where   
there are two groups of shoes - one which is cheap quality but looks nice and   
carries a high price tag or one that is a higher quality but does not look as   
nice but carries a slightly lower price tag.


